Maintenance planning on french military aircraft operations

### 2017-11-13 latex/ROADEF/article_roadef_2018.tex
added R scripts to generate calendar, input data with calendar in Excel and small aesthetic modifications in latex

% Exemple d'utilisation de la classe roadef2018 pour le congrès ROADEF 2018 (http://roadef2018.labsticc.fr)

@@

% \usepackage{fontspec}

@@

% Les AUTeurs et leur numéro d'affiliation
\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1} }

% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
\institute{
ISAE... \\
\email{\{franco.peschiera\}@isae-supaero.fr}
}

% Création de la page de titre
\maketitle
\thispagestyle{empty}

% Les mots-clés
\keywords{optimization, planning, military, maintenance}

@@

    \begin{tabular}{p{30mm}lp{90mm}}
        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ (including t=0) \\  
        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ (including t=0) \\  
        $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
        $start_{at}$    & : & 1 if resource $a$ starts maintenance in period $t$ \\
        $available$     & : & minimum number of available resources in any period \\
        $maintenance$   & : & maxium number of resources in maintenance in any period \\
    \end{tabular}
    \bigskip

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN VARS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
                \hspace{10mm} t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
                \hspace{10mm} t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % BALANCES
        %%%%%%%%%%%%%%%%%%%
        % definition of consumed resource hours:
        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
                \hspace{10mm} a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
        % remaining used time
        & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
        % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
        & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % STATES
        %%%%%%%%%%%%%%%%%%%
        % if start maintenance, at least X periods
        & state_{ast_2} \geq start_{at_1} \notag\\
                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start1}\\
        % if maintenance doesn't start, we cannot have maintenance:
        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start2}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt}
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'V' \label{eq:state1}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % OBJECTIVE
        %%%%%%%%%%%%%%%%%%%
        % min available
        & available \leq \sum_{a} state_{ast}
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:available}\\
        % max maintenance
        & maintenance \geq \sum_{a} state_{ast}
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:maintenance}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % HORIZON
        %%%%%%%%%%%%%%%%%%%
        % TODO: initialize maintenance states too!
        % rut initialize
        & rut_{at} = RUT\_INI_a
                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
        % ret initialize
        & ret_{at} = RET\_INI_a
                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
        % fixed states during planning period
        & state_{ast} = STATE\_INI_{at}
                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \{\} \label{eq:fixedstates}\\
    \end{align}

@@

% Version "on-line" de la bibliographie, mais il est
% également possible d'utiliser un fichier ".bib" et d'utiliser BibTeX

>>>>

@@

% \usepackage{fontspec}
% This command is to use simple quotes inside math expressions:
\newcommand{\mq}[1] {`#1\textrm'}

@@

\author{Franco Peschiera, Alain Haït, Olga Battaïa }

\institute{
ISAE... \\
\email{\{franco.peschiera\}@isae-supaero.fr}
}

\maketitle
\thispagestyle{empty}

\keywords{optimization, planning, military, maintenance}

@@

    \begin{tabular}{p{30mm}lp{90mm}}
        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ (including t=0) \\  
        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ (including t=0) \\  
        $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
        $start_{at}$    & : & 1 if resource $a$ starts maintenance in period $t$ \\
        $available$     & : & minimum number of available resources in any period \\
        $maintenance$   & : & maximum number of resources in maintenance in any period \\
    \end{tabular}
    \bigskip

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN VARS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
                & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
                & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % BALANCES
        %%%%%%%%%%%%%%%%%%%
        % definition of consumed resource hours:
        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
                & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
        % remaining used time
        & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
        % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
        & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % STATES
        %%%%%%%%%%%%%%%%%%%
        % if start maintenance, at least X periods
        & state_{ast_2} \geq start_{at_1} \notag\\
                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
        % if maintenance doesn't start, we cannot have maintenance:
        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt} \notag\\
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1 \notag\\
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % OBJECTIVE
        %%%%%%%%%%%%%%%%%%%
        % min available
        & available \leq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
        % max maintenance
        & maintenance \geq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % HORIZON
        %%%%%%%%%%%%%%%%%%%
        % TODO: initialize maintenance states too!
        % rut initialize
        & rut_{at} = RUT\_INI_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
        % ret initialize
        & ret_{at} = RET\_INI_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
        % fixed states during planning period
        & state_{ast} = (STATE\_INI_{at} = s)
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \mq{} \label{eq:fixedstates}\\
    \end{align}

@@
